---
title: "PGS-Catalog"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
author: pgs-catalog-harmoniser
params:
  logs_directory: "../tests/input/logs"
  meta_file: "../tests/input/pgs_all_metadata.xlsx"
  output_filename: "output"
---

```{r setup, include=FALSE}

library(readr)
library(stringr)
library(readxl)
library(ggplot2)
library(DT)
library(dplyr)
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
meta <- read_excel(params$meta_file, sheet = "Readme", col_names = FALSE)
datatable(meta)
```

## Builds


```{r}
scores <- read_excel(params$meta_file, sheet = "Scores") %>%
    mutate(
      original_build = case_when(
        `Original Genome Build` == "hg19" | `Original Genome Build` =="GRCh37"  ~ "hg19",
        `Original Genome Build` == "hg38" | `Original Genome Build` =="GRCh38"  ~ "hg38",
        `Original Genome Build` == "NR" ~ "NR",
        TRUE ~ "Other",
      )
  )

ggplot(scores, aes(x = original_build)) +  geom_bar()

scores %>% group_by(original_build) %>% summarize(count = n()) %>% mutate(percentage = paste(round(count / sum(count) * 100, 2), "%")) %>% arrange(desc(percentage)) %>% datatable()

other_build <- scores %>% filter(original_build == "Other")


```

## Variants

```{r}
ggplot(scores, aes(x = `Number of Variants`)) +   geom_histogram()
```


```{r}

# Load all log files
result <- data.frame() %>%
  mutate(
    score = NA,
    format = NA,
    original_variants = NA,
    variants = NA,
    coverage = NA,
    chain_file = NA,
    dbsnp_file = NA,
    error = NA
  )

files <- list.files(path=params$logs_directory, pattern="*.log", full.names=TRUE, recursive=FALSE)
for ( filename in files){
  text <- read_lines(filename)
  text <- paste(text, collapse='\n')
  name <- str_match(text, "Input File: (.*).original.txt.gz")
  format <- str_match(text, "Input File Format: (.*)")
  variants <- str_match(text, "Number Variants Output: (.*)/(.*)")
  chain_file <- str_match(text, "Liftover using chain file '(.*)'")
  dbsnp_file <- str_match(text, "Resolve rsIDs using index file '(.*)'")
  error <- str_match(text, "\\*\\*\\* ERROR \\*\\*\\*  (.*)");

  result[nrow(result) + 1,] <- c(name[2], format[2], strtoi(variants[3]), strtoi(variants[2]), round(strtoi(variants[2]) /strtoi(variants[3]),2), chain_file[2], dbsnp_file[2], error[2])
}

result <- merge(result, scores, by.x  ="score", by.y = "Polygenic Score (PGS) ID")

result <- result %>%
  mutate(
    has_error = case_when(
      is.na(error) ~ TRUE,
      TRUE ~  FALSE
    )
  )


# TODO: build table
```

## Unreadable Files

```{r}

ggplot(result, aes(x = has_error)) +  geom_bar()
result %>%
  group_by(has_error) %>%
  summarize(count = n()) %>%
  mutate(percentage = paste(round(count / sum(count) * 100, 2), "%")) %>%     
  arrange(desc(percentage)) %>%
  datatable()


errors <- result %>%
  filter(!is.na(error)) %>%
  select(score, error)
errors %>% datatable()

```

## Format

```{r}
ggplot(result, aes(x = format)) +  geom_bar()
result %>%
  group_by(format) %>%
  summarize(count = n()) %>%
  mutate(percentage = paste(round(count / sum(count) * 100, 2), "%")) %>%
  arrange(desc(percentage)) %>%
  datatable()
```

## Coverage

```{r}
ggplot(result, aes(x=as.double(coverage))) +  geom_histogram()
```


## Low Coverage

```{r}
coverage_filter = 0.75
low_coverage <- result %>% filter(coverage < coverage_filter)
low_coverage %>%
  select(score, coverage, variants, original_variants, format, original_build) %>%
  datatable()
```


## Final Scores

```{r}
final <- result %>%
  filter(coverage >= coverage_filter & is.na(error)) %>%
  mutate (
    action = case_when(
      !is.na(chain_file) ~ "LIFT_OVER",
      !is.na(dbsnp_file) ~ "RSID_TO_COORDINATES",
      TRUE ~ "NONE"
    )
  )

ggplot(final, aes(x = action)) +  geom_bar()

final %>%
  group_by(action) %>%
  summarize(count = n()) %>%
  mutate(percentage = paste(round(count / sum(count) * 100, 2), "%")) %>%
  arrange(desc(percentage)) %>%
  datatable()

final %>%
  select(score, coverage, variants, original_variants, format, original_build, action) %>%
  datatable()

final %>%
   select(score, coverage, variants, original_variants, format, original_build, action) %>%
   write.csv(file = params$output_filename,  row.names = FALSE)
```

## Summary

- Input: `r nrow(scores)` scores
- Unsupported Build: `r nrow(other_build)` ( `r round(nrow(other_build) / nrow(scores),2)`%)
- Unreadable Files: `r nrow(errors)` ( `r round(nrow(errors) / nrow(scores),2)`%)
- Coverage < `r coverage_filter`: `r nrow(low_coverage)` ( `r round(nrow(low_coverage) / nrow(scores),2)` %)
- Output: `r nrow(final)` scores (`r round(nrow(final) / nrow(scores),2)`%)
